---
format: 
  html:
    mainfont: "Courier New"
    fontsize: 10pt
    output-file: "index.html"
editor: visual
---

::: content-hidden
## loading packages

```{r}
library(tidyverse)
library(lubridate)
library(knitr)
library(readxl)
library(flextable) #using this for a powerpoint output
library(magrittr) # for the flextable theme
library(officer)

#change again5
```

# 1. import agenda

Change to how you have this

```{r}
schedule_raw <- read_excel("C://Users//ltee4799//OneDrive - The University of Sydney (Staff)//Matilda Staff - Stats and Coding Collective//Schedule//Ideas, resources, and agenda.xlsx", sheet = "Agenda")
```

## 1.1. convert the dates to something we can use

excel does dates in a weird way. this converts them to dates that r can read safely

```{r}
event_dates <- schedule_raw %>% mutate( 
  Date = ymd(Date) #ymd as that's what excel did (weirdly)
)
```

# 2. build calendar data

we currently don't have the dates for mondays and we don't note which mondays and thursdays fall in the same week. So this section creates a table where we have the date, weekenday, and what week it's in. Only for mondays and thursdays though bc we don't need those other days.

- it just takes the min and max date from the excel spreadsheet agenda (through event_dates), nothing else.

- **here may be a good spot to restrict to current or future weeks, rather than indefinitely**

```{r}
#specify this week so that we can filter to only this week
this_monday <- floor_date(Sys.Date(), "week", week_start = 1)

#make a calendar table
calendar_tbl <- tibble(
  Date = seq(
    floor_date(min(event_dates$Date), "week", week_start = 1),
    ceiling_date(max(event_dates$Date), "week", week_start = 1) - days(1),
    by = "day"
  )
) %>%
  filter(Date >= this_monday) %>% #restrict to current and future week only
  filter(wday(Date, week_start = 1) %in% c(1, 4)) %>%   # Mon + Thu
  mutate(
    weekday = wday(Date, label = TRUE, week_start = 1),
    week = floor_date(Date, "week", week_start = 1)
  )
```

# 3. combine events + calendar information

## 3.1. combine information

add the actual information in the agenda from 'event_dates' to 'calendar_table'. This code makes a table that has all the information that we want, except the colouring :)

```{r}
combined_tbl <- calendar_tbl %>%
  left_join(event_dates, by = "Date") %>%  # join Excel events onto the calendar structure
  mutate(
    topic = case_when(
      weekday == "Mon" ~ "Stats drop-in", #this adds in that stats drop ins occur on mondays
      weekday == "Thu" ~ topic #this adds in that whatever topic we indicated in the excel spreadsheet gets popped in here
    )
    ) %>%
  mutate (
    type = case_when(
      weekday == "Mon" ~ "drop-in", #changing the type of all mondays to 'drop-in' so that colour adding will be easier :)
      weekday == "Thu" ~ type #adding in type from excel spreadsheet
    )
    ) %>% 
  mutate(
    weekday = as.character(weekday), #changing the sort of data the weekday is so it works better below
    label_ppt = paste( #new collumn with the following
      paste(weekday, format(Date, "%d %B")), #first 'row' being dates
      paste("TOPIC:", topic), #then topic
      ifelse(
        is.na(person_leading),
        "", #bc R cannot return something like 'NULL' with zero length
        paste("LEADER/s:", person_leading)
      ),
      sep = "\n" #all separated by a new line
      )
  )
```

## 3.2. make it look like a calendar

take combined_tbl and make it look like a calendar, rather than a table (i.e., have two rows, one for monday, and one for thursdays, and the information needed underneath)

```{r}
calendar_wide <- combined_tbl %>%
  select(week, weekday, label_ppt) %>%
  pivot_wider(
    names_from = weekday,
    values_from = label_ppt) %>%
  select(Mon, Thu)
```

## 3.3. add colours

identifying the colours

```{r}
  fill_colours <- c(
  "presentation" = "#f5e68e",  # yellow
  "work through" = "#85d184",  # light green
  "drop-in"      = "#E6E6E6",  # grey
  "TBD"          = "#f58e8e",   # red / pink
  "NA"           = "BDBDBD"  # grey
  )
  

```

making a lookup table for colours that has the same format as 'calendar_wide'

```{r}
calendar_fill <- combined_tbl %>% #change 'combined _tbl' to do this
  mutate(weekday = as.character(weekday)) %>% #read weekday as characters
  select(week, weekday, type) %>%
  pivot_wider(
    names_from = weekday,
    values_from = type
  ) %>%
  select(Mon, Thu) #so that it also doesn't include weekday as well and the rest down't get mixed up
```

# 4. create colours flextable

theme design from (<https://www.ardata.fr/en/flextable-gallery/2020-12-01-creating-theme-function/>)

```{r}
theme_design <- function(x) {
  x <- border_remove(x)
  std_border <- fp_border_default(width = 4, color = "white")
  x <- fontsize(x, size = 10, part = "all")
  x <- font(x, fontname = "Courier New", part = "all")
  x <- align(x, align = "center", part = "all")
  x <- bold(x, bold = TRUE, part = "header")
  x <- color(x, color = "black", part = "all")
  x <- padding(x, padding = 6, part = "all")
  x <- border_outer(x, part="all", border = std_border )
  x <- border_inner_h(x, border = std_border, part="all")
  x <- border_inner_v(x, border = std_border, part="all")
  x <- set_table_properties(x, layout = "fixed")
  x
}
```

making it look nice

```{r}
ft <- flextable(calendar_wide) %>%
  set_header_labels(
    Mon = "STATS ON MONDAYS", 
    Thu = "CODING ON THURSDAYS"
  ) %>%
  bg(part = "header", bg= "#4ea66d") %>% #green for header
  theme_design()
  

#adding basic formatting
ft <- ft %>%
  align(align = "left", part = "all") %>% #align of content
  valign(valign = "top", part = "all") %>% #align of values, i think, in the cells
  align(align = "center", part = "header") %>% #to keep header centred
  autofit()

#setting width

ft <- ft %>%
  width(j = "Mon", width = 3.0) %>%   # widen Monday column
  width(j = "Thu", width = 3.0) %>%   # optional: adjust Thursday
  align(align = "left", part = "all") %>%
  valign(valign = "top", part = "all") %>%
  align(align = "center", part = "header")
  

#apply background colours cell-bycell
  for (i in seq_len(nrow(calendar_fill))) {
    for (j in seq_len(ncol(calendar_fill))) {
      type_ij <- calendar_fill[i, j][[1]]
  
      bg_colour <- if (is.na(type_ij)) {
        fill_colours[["NA"]]
      } else {
        fill_colours[[as.character(type_ij)]]
      }
  
      ft <- bg(ft, i = i, j = j, bg = bg_colour)
    }
  }


  
#making a key flextable for the colours
  
  key_df <- data.frame(
    Key = names(fill_colours)[1:4]
)
#make the flextable
  key_ft <- flextable(key_df) %>%
    bg(i = 1, bg = fill_colours["presentation"]) %>%
    bg(i= 2, bg=fill_colours["work through"]) %>%
    bg(i=3, bg = fill_colours["drop-in"]) %>%
    bg(i=4, bg = fill_colours["TBD"]) %>% 
    theme_design() %>%
    autofit()
```

theme design (from

# 5. table
:::

```{r}
#| echo: false
#| results: asis
cat("**Stats and Code Collective Agenda**\n\n")
cat(last_run <- paste("Last updated:", format(Sys.time(), "%d %B %Y %H:%M\n\n")))
cat("On Mondays and Thursdays at 1pm")

ft 

key_ft

cat("a work through is where we follow a tutorial from the internet to learn together (e.g., ggplot) or someone talks through how they use a coding feature") 
```

::: content-hidden
# 6. feedback to integrate

- all cap bold monday and thursday at top and full words

- darker dates

- monday longer so doesn't look like it's the date of thursday - EQUAL width of column

- bolds

- making clear what type of thing is what colour (e.g., PRESENTATION:.... or WORK THROUGH...) within the thing itself

- make dates clearer (maybe darker colour/grey

- title (obvs)

- people given PRESENTER in front

- 

\`\`\`
:::
